---
title: "Peer Assessment 1"
author: "Kumar Aiyer"
date: "Friday, August 15, 2014"
output: html_document
---

## Loading Data and Cleaning

We load the data after dowloading to the data directory. The data loading need not be repeated each time the document is _knitted_ .

```{r dataloding, echo=TRUE}

activity.raw <- read.csv("data/activity.csv")
activity.nona.df <- activity.raw[!is.na(activity.raw$steps),]
```

## Mean and Median Total Number of Steps per Day

Next, we compute the total number of steps per day. To compute this, we first filter out the NAs and then compute the number of steps per day. We then plot a histogram of the data and report the mean and median.

```{r stepsperday, echo=TRUE}
total_steps_per_day <- aggregate(activity.nona.df$steps, by=list(activity.nona.df$date), FUN=sum)$x
hist(total_steps_per_day)
mean(total_steps_per_day)
median(total_steps_per_day)

```

## Average Daily Activity Pattern

We than aggregate the data by 5min intervals across multiple days, plot the time series and find the interval with the maximum steps

```{r dailypattern,echo=TRUE}
mean_steps_per_5min <- aggregate(activity.nona.df$steps, by=list(activity.nona.df$interval), FUN=mean)
names(mean_steps_per_5min) <- c("interval", "steps")
plot(mean_steps_per_5min$interval, mean_steps_per_5min$steps, type='l')
mean_steps_per_5min[mean_steps_per_5min$steps==max(mean_steps_per_5min$steps),]
```
## Imputing Missing Values

In this we find the number of missing values and then fill the missing NA with the average across daily values for the particular interval. 

```{r missingvalues,echo=TRUE}
sum(is.na(activity.raw$steps)==TRUE)
mean_steps_per_5min <- aggregate(activity.nona.df$steps, by=list(activity.nona.df$interval), FUN=mean)
names(mean_steps_per_5min) <- c("interval", "steps")
activity.new.df <- activity.raw
for (i in dim(activity.raw)[1]) {
  if (is.na(activity.new.df[i,]$steps)) {
    new_steps <- mean_steps_per_5min[mean_steps_per_5min$interval==activity.new.df[i,]$interval,]$steps
    new_steps
    activity.new.df[i,]$steps <- new_steps
  }
}
total_steps_per_day <- aggregate(activity.new.df$steps, by=list(activity.new.df$date), FUN=sum)$x
hist(total_steps_per_day)
mean(total_steps_per_day)
median(total_steps_per_day)
```